\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-15 arch.tex
rearranging text from concepts into europa-basics

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as

  each reactor declares its planning look-ahead $\pi$ and expected
  latency $\lambda$
  % \comment{$\pi \& \lambda$ both defined here for the first time}
  before being able to produce plan to the agent depending on where it
  is located in a reactor dependency graph such as shown in
  Fig. \ref{fig:agent}.

>>>>

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as
  each reactor declares its planning look-ahead $\pi$ and expected
  latency $\lambda$ before being able to produce plan to the agent
  depending on where it is located in a reactor dependency graph such
  as shown in Fig. \ref{fig:agent}.

### 2012-02-15 concepts.tex
rearranging text from concepts into europa-basics

\section{Foundational Concepts}
\label{sec:concepts}

\texttt{T-REX} is an adaptive, artificial intelligence based
controller and provides general framework for building reasoning
systems for real-world autonomous vehicles. At MBARI \texttt{T-REX} is
used for AUV control; another instantiation of the system is being
used for control of a terrestrial personal robot \cite{pr2,
  Meeussen:2010dn}. The development of \texttt{T-REX} has been
targeted at surveying a number of oceanographic features which are
dynamic and unpredictable spatio-temporally. Typically this requires
our AUVs to balance the goals of coverage spatially while
opportunistically following features of scientific interest and to do
so while being operationally aware of its own limitations in terms of
resources (typically the battery state of charge) and overall
proximity to other observational assets for obtaining scientific
ground-truth. We are therefore interested in representational
frameworks that allow robots to pursue long-term objectives while
choosing short-term gain and can gracefully deal with exogenous or
endogenous change.

To enable this responsiveness to external observations, the agent has
to be able to synthesize plans insitu and to re-plan. We use a
temporal constraint-based planner \eu with a demonstrated legacy of
having flown on NASA space missions \cite{jonsson00,bresina05,
  barreiro09}. Our autonomy architecture brings three key innovations
for AUV adaptation: the use of flexible plan representations,
compositional control with the use of partitioned networks and
off-line learning to inform insitu environmental state estimation. In
this section we detail some of the key concepts relevant to flexible
plan representation. Section \ref{sec:arch} has details the
\rx partitioned architecture and estimation driven vehicle
adaptation.

\subsection{The basics}
\label{sec:basics}

\eu uses a \emph{domain model} written in a declarative language
(NDDL: New Domain Description Language), together with initial
conditions and goals also specified in NDDL, to construct a set of
temporal relations that must be true at the start time. These models
include assertions about the physics of the vehicle, i.e how it
responds to external stimulus and internally driven goals. By
propagating these relations forward using Simple Temporal Networks
\cite{dechter91} and applying goal constraints, \eu can select a set
of conditions that should be true in the future, where some of these
conditions will correspond to actions the agent must take. The planner
can backtrack and try another path during search if a goal cannot be
reached while being capable of discarding unachievable goals.

Traditionally robot execution has relied on dispatching commands at
precise times. Such linear sequences of precisely timed commands give
no ability to adjust execution on the basis of sensory
information. Although some commands can issue tests on sensor
readings, these tests have the objective of verifying whether expected
execution conditions are occuring. If not, the state of the system is
declared off-nominal and execution of the sequence of commands is
interrupted. More recently, executives have been proposed and
implemented that significantly broaden the way robots can be commanded
\cite{mus98,alami:1998p820}. For example, the Remote Agent executive
interpreted a \textit{temporally flexible plan} which represents each
start time as a variable and contains an explicit network of bounded
delay constraints between such variables.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{figs/flexible-timelines.pdf}
\caption{\small Tokens with flexible temporal intervals and parametric
  constraints between tokens. This example shows the triggering of a
  water sampler based on a feature threshold while the vehicle
  Yo-Yo's. The \texttt{Waypoint\_Yo-Yo} token has a flexible duration,
  start \& end times.}
\label{fig:flex-timelines}
\vskip+0.1cm
\end{figure}

Unlike a traditional fixed time-tagged command sequences, such
flexible plans leave room for adaptation at execution time. When the
executive considers when to start a task, it propagates information
through the constraint network, computes a time bound for the
variable, selects an actual execution time within the bound, and
starts the task at that time. Temporally flexible plans therefore,
express a \textit{range of possible outcomes} of the robots
interaction with the environment within which the executive can elect
at run time the most appropriate one for the actual execution
conditions. The fact that constraints are explicitly represented
ensures that through constraint propagation the executive will respect
global limits expressed in the plan (e.g., don't start a task until a
certain condition has been satisfied but still satisfying some global
deadline). Such flexibility is critical when dealing with dynamic
ocean conditions where precise timing of a robotic action might be
indeterminate. Further, the advantage of flexibility can be contrasted
with the consequences of the intrinsic inflexibiliy of traditional
command sequences. Because they are inflexible, sequences must
necessarily be designed considering worst case scenarios.

\eu uses a \emph{state variable} representation to describe the
evolution of state over time. The instantiated history of such state
variable evolution over a temporal horizon we call \emph{timelines}
and which represent a single thread in the execution of a concurrent
system. At any given time each thread can execute a single procedure.
Thus each timeline consists of a sequence of procedures which
encapsulate and describe state evolution; we call these instantiated
atomic entities \emph{tokens}.  A token therefore describes a
procedure invocation, the state variables on which it can occur, the
parameter values of the procedure, and the time values defining the
interval. We allow encapsulation of uncertainty within these tokens
with a range of start and end times and parameters, all of which are
encoded as variables. A constraint solver in turn manipulates these
variables defined in a \eu domain model. For example, consider
a scientific need to take a water samples 100 meters from a hotspot
while an AUV is performing a Yo-Yo in the water-column. Two samples
are needed if the feature's signal is above a threshold; one
otherwise. The token that is capturing the sensory threshold has a
parametric constraints to the token which fires the requisite water
sampler. In addition, the start time of the water sampling procedure
is highly dependant on the variability of sub-sea currents and actual
speed of the vehicle. Therefore a number of values are possible for
the start times and duration of the sampling all of which are valid
combinations for desired outcomes.

\begin{figure}
\centering
\includegraphics[scale=0.3]{figs/Allen-algebra.pdf}
\caption{\small Temporal relations defined within the planner are
  based on \texttt{Allen Algebra} relations shown above.}
\label{fig:allen-algebra}
\vskip-0.3cm
\end{figure}

@@

\item[\textbf{Variables}] Values that need to be represented to
  describe the problem domain and over which we may want to specify
  constraints. In the Shopping Agent example, the times at which the
  agent needs to leave or be back, or executes a purchase, are all
  instances where variables representing time would be used. 
  As explained in the CP section, a variable can take values from a domain. In \eu, domains are defined over a specific data type, the primitive data types supported are:
  \begin{enumerate}
    \item \textit{String}: sequences of characters, like "Red", "Spirit", etc 
    \item \textit{Boolean}: {True,False}
    \item \textit{Numeric}: integers and floats
    \item \textit{Object}: reference to an object instance (see below for a description of objects in \eu)
  \end {enumerate}
  Two types of variable Domains can be defined over a data type: 
  \begin{enumerate}
    \item \textit{Enumerated}: Finite sets over any data type, specified explicitly, for instance: ["Red","Yellow","Green"], [1,18,32], etc
    \item \textit{Interval}: Only for Numeric data types, they are specified by a [LowerBound,UpperBound] pair, for instance: [1,10], [5.0,100.0], etc.
  \end {enumerate}
  \eu's ability to represent and reason over numeric intervals is a very useful characteristic that allow users to create flexible plans. \comment{say more about the benefits of flexibility?}

@@

\begin{enumerate}
    \item \textit{Timelines}: Often objects in a domain must be described by exactly one
    token for every given timepoint in the plan. Any instances of a class derived from \eu's built-in
    Timeline class will induce ordering requirements among its tokens
     to ensure no temporal overlap may occur among them  \cite{mus94}. 
     \comment{add small example?}

@@

\begin{enumerate}
    \item If there are any tokens in the plan that are compatible with the new goal, \eu may attempt to satisfy the goal by merging it with any of them (if there is more than one it can try them one at a time), in this case, the state of the token would change to MERGED.
    \item Instead of merging with existing tokens in the plan, \eu may decide to insert the new token into the plan, in that case its state would change to ACTIVE.
\end{enumerate}

@@

\begin{enumerate}
   \item \textbf{Domains, Variables and Objects} to represent the entities (Objects) and attributes (Variables) needed to represent a problem domain	
    \item \textbf{Tokens} Temporally scoped predicates to represent states and actions in time. 
    \item \textbf{Token States} which are necessary for the implementation of planning operations on a partial plan.
    \item \textbf{Constraints} to describe relationships among attributes of objects and tokens. \eu has a built-in library of constraints that covers time, resources and others that are common in real life problem domains.
    \item \textbf{Domain Rules} to describe internal and external relationships on and
     between tokens respectively. Rules are applied to active tokens (master tokens) 
     referred to as masters and typically produce inactive tokens (slaves) which are a key vehicle through which the planning
     process evolves. \comment{citation, or perhaps remove this statement?:}The rule structure in \eu differs from the more restrictive commitment to preconditions and effects in classical
     planning which prohibits durative actions and disembodied effects (i.e. effects which may not occur until some temporal distance after
     the end of the action).    
    \item \textbf{Timelines and Resources} are built in implementations of Object types that appear frequently in real life domains.
\end{enumerate}

>>>>

@@

\item[\textbf{Variables}] Values that need to be represented to
  describe the problem domain and over which we may want to specify
  constraints. In the Shopping Agent example, the times at which the
  agent needs to leave or be back, or executes a purchase, are all
  instances where variables representing time would be used.  As
  explained in the \textsf{CP} section, a variable can take values
  from a domain. In \eu, domains are defined over a specific data
  type, the primitive data types supported are:

  \begin{enumerate}
  \item \textit{String}: sequences of characters, like "Red",
    "Spirit", etc

  \item \textit{Boolean}: {True,False}

  \item \textit{Numeric}: integers and floats

  \item \textit{Object}: reference to an object instance (see below
    for a description of objects in \eu)

\end {enumerate}

Two types of variable Domains can be defined over a data type:

  \begin{enumerate}
  \item \textit{Enumerated}: Finite sets over any data type, specified
    explicitly, for instance: ["Red","Yellow","Green"], [1,18,32], etc
  \item \textit{Interval}: Only for Numeric data types, they are
    specified by a [LowerBound,UpperBound] pair, for instance: [1,10],
    [5.0,100.0], etc.
  \end {enumerate}

  \eu's ability to represent and reason over numeric intervals is a
  very useful characteristic that allow users to create flexible
  plans. \comment{say more about the benefits of flexibility?}

@@

\item \textit{Timelines}: Often objects in a domain must be described
  by exactly one token for every given timepoint in the plan. Any
  instances of a class derived from \eu's built-in Timeline class will
  induce ordering requirements among its tokens to ensure no temporal
  overlap may occur among them \cite{mus94}.  \comment{add small
    example?}

@@

\item If there are any tokens in the plan that are compatible with the
  new goal, \eu may attempt to satisfy the goal by merging it with any
  of them (if there is more than one it can try them one at a time),
  in this case, the state of the token would change to MERGED.

\item Instead of merging with existing tokens in the plan, \eu may
  decide to insert the new token into the plan, in that case its state
  would change to ACTIVE.

@@

\begin{enumerate}
\item \textbf{Domains, Variables and Objects} to represent the
  entities (Objects) and attributes (Variables) needed to represent a
  problem domain

\item \textbf{Tokens} Temporally scoped predicates to represent states
  and actions in time.

\item \textbf{Token States} which are necessary for the implementation
  of planning operations on a partial plan.

\item \textbf{Constraints} to describe relationships among attributes
  of objects and tokens. \eu has a built-in library of constraints
  that covers time, resources and others that are common in real life
  problem domains.

\item \textbf{Domain Rules} to describe internal and external
  relationships on and between tokens respectively. Rules are applied
  to active tokens (master tokens) referred to as masters and
  typically produce inactive tokens (slaves) which are a key vehicle
  through which the planning process evolves. \comment{citation, or
    perhaps remove this statement?:}The rule structure in \eu differs
  from the more restrictive commitment to preconditions and effects in
  classical planning which prohibits durative actions and disembodied
  effects (i.e. effects which may not occur until some temporal
  distance after the end of the action).

\item \textbf{Timelines and Resources} are built in implementations of
  Object types that appear frequently in real life domains.
